### 2011-12-22 GUIErrorDetector/paper/abstract.tex
tweak

In this paper, we present a general technique to find such \textit{invalid thread access}
errors in multithreaded GUI applications. We formulate finding invalid
thread access errors as a call graph
reachability problem. Our technique starts from call graph entry nodes and statically explores
all possible paths that may cause a non-UI thread to access a GUI object.
However, the heavy use of reflection in modern GUI frameworks makes existing
call graph construction algorithms insufficent to build a both complete and precise
call graph. To overcome such limitations, we present a combined RTA and k-CFA
call graph analysis (that XXX). % to overcome such limitations.% of analyzing GUI applications.

>>>>

In this paper, we present a general technique to find such \textit{invalid thread access}
errors in multithreaded GUI applications. We formulate finding invalid
thread access errors as a call graph
reachability problem. Our technique starts from call graph entry nodes and statically explores
all possible paths that may cause a non-UI thread to access a GUI object.
However, the heavy use of reflection in modern GUI frameworks makes existing
call graph construction algorithms insufficient to build a both complete and precise
call graph. To overcome such limitations, we present a combined RTA and k-CFA
call graph analysis (that XXX). % to overcome such limitations.% of analyzing GUI applications.

### 2011-12-22 GUIErrorDetector/paper/introduction.tex
tweak

The \textit{event dispatching thread}, or called \textit{UI thread}, is a global special
thread initilized by the GUI framework, where all event-handling code
is executed. All code that interacts with GUI objects must also
execute on this thread.  There are several advantages to adapt the single-thread rule:

@@

\item GUI developers do not have to have an in-depth
unerstanding of concurrency programming. A GUI framework in which all components
must fully support multithreaded access can be difficult to extend, particularly
for developers who are not expert at threads programming.

\item Events are dispatched in a predictable order. Mouse and keyboard events, timer events, 
paint requests, and other user-defined events are dispatched from the same event queue.
In constrast, in a GUI framework without adapting the single-thread rule where components
support multithreaded access, any component changes can be interleaved with event
processing at the whim of the thread scheduler. This makes program behavior comprehension hard,
 and comprehensive testing difficult or impossible.

\item Single-thread rule incurs less overhead.
A GUI framework that attempts to carefully lock critical sections can spend a substantial
amount of time and space managing locks. Whenever the framework calls a method that might
be implemented in client code (for example, any public or protected method in a public class),
the framework must save its state and release all locks so that the client code can grab locks
if necessary. When GUI objects return from the method, the framework must regrab their locks and
restore states. Therefore, all applications using the framework will bear the cost of this, even though most
applications do not require concurrent access to the GUI.

@@

In a multithreaded GUI application, the single-thread rule can be
easily violated since a spawned non-UI thread often needs to update
the GUI to display results after its computation task is finished.
An \textit{invalid thread access} error occurs when
a GUI object is accidentally accessed by a non-UI thread. In practice, such
invalid thread access error is \textit{frequent}, \textit{pervasive}, \textit{serverse}.
and \textit{difficult to avoid}.

@@

Software testing, the most widely used method to ensure program correctness,
is often insufficient to detect such errors. That is because the space of possible interactions
with a GUI is enormous, in that each sequence of GUI events can result in
a difficult state, and each GUI event may need to be evaluated in all of
these states. It is almost impossible for testing to cover every possible
states in practice. In fact, for most reported bugs in eclipse, only very specific
UI interaction sequences can trigger the error. A large software system like eclipse
has a well-designed test suite that achieves failry high coverage; but many
 bug-triggering UI sequences are still not covered.

Program refactoring~\cite{Mens:2004}, another popular way to improve software quality,
is also inadaqute to eliminate this error.  A standard way to fix such
invalid thread access error is to replace direct GUI object accessing operations
with asynchronized message passing, to ensure GUI object are accessed in
the UI thread. For example, fixing the bug in Figure~\ref{fig:androiderror}
needs to wrap line 222 inside a \CodeIn{post} message passing method\footnote{The
\CodeIn{post} method is in class \CodeIn{android.widget.View}, providing
a standard utility to send messages to the UI-thread}, as follows:

@@

One may wonder that a straightforward program transformation to wrap
every possible UI-accessing operation with asynchronized messsage passing
may solve this problem. However, such an approach is problematic,
because using asynchronized message passing has no timing
gurantee. It is entirely possible that, in some cases, a GUI object has already been
disposed before the message sent to arrives. Thus, programmers can not use
it everywhere; instead, they must choose carefully to put it in appropriate places
with extreme caution.

@@

In this paper, we choose to use static analysis to find such potential errors.
Static analysis has several advantages while comparing to dynamic approaches like testing for this problem.
First, a static analysis can explore every path of the program without executing the code, and
thus there is no need to construct non-trivial tests and harnesses.
Second, a static analysis can soundly find potentially errors: if it
reports no warnings, the code is guranteed to be bug-free.

@@

\item \textbf{Technique.} We present a general technique to find
such errors. In particular, to deal with the unique features in
multithreaded GUI programs, we propose a coombined RTA and k-CFA
call graph construction algorithm

\item \textbf{Implemetnation.} We instantiate our technique for four
popular GUI platforms: SWT, eclipse plugin, Swing, and Android. Our
tool implementation is publicly available at:
\url{http://guierrordetector.googlecode.com}

>>>>

The \textit{event dispatching thread}, or called \textit{UI thread}, is a global special
thread initialized by the GUI framework, where all event-handling code
is executed. All code that interacts with GUI objects must also
execute on this thread.  There are several advantages to adapt the single-thread rule:

@@

\item GUI developers do not have to have an in-depth
understanding of concurrency programming. A GUI framework in which all components
must fully support multithreaded access can be difficult to extend, particularly
for developers who are not expert at threads programming.

\item Events are dispatched in a predictable order. Mouse and keyboard events, timer events, 
paint requests, and other user-defined events are dispatched from the same event queue.
In contrast, in a GUI framework without adapting the single-thread rule where components
support multithreaded access, any component changes can be interleaved with event
processing at the whim of the thread scheduler. This makes program behavior comprehension hard,
 and comprehensive testing difficult or impossible.

\item Single-thread rule incurs less overhead.
A GUI framework that attempts to carefully lock critical sections can spend a substantial
amount of time and space managing locks. Whenever the framework calls a method that might
be implemented in client code (for example, any public or protected method in a public class),
the framework must save its state and release all locks so that the client code can grab locks
if necessary. When GUI objects return from the method, the framework must re-grab their locks and
restore states. Therefore, all applications using the framework will bear the cost of this, even though most
applications do not require concurrent access to the GUI.

@@

In a multithreaded GUI application, the single-thread rule can be
easily violated since a spawned non-UI thread often needs to update
the GUI to display results after its computation task is finished.
For example, a server program can get requests from other programs
that might be running on different machines. These requests can come at any time,
and they result in one of the server's methods being invoked in some thread.
After handling the request, the GUI-update code must be executed on the UI-thread.
If not, an \textit{invalid thread access} will error occur. In practice, such
invalid thread access error is \textit{frequent}, \textit{pervasive}, \textit{severe}.
and \textit{difficult to avoid}.

@@

Software testing, the most widely used method to ensure program correctness,
is often insufficient to detect such errors. That is because the space of possible interactions
with a GUI is enormous, in that each sequence of GUI events can result in
a difficult state, and each GUI event may need to be evaluated in all of
these states. It is almost impossible for testing to cover every possible
states in practice. In fact, for most reported bugs in eclipse, only very specific
UI interaction sequences can trigger the error. A large software system like eclipse
has a well-designed test suite that achieves fairly high coverage; but many
 bug-triggering UI sequences are still not covered.

Program refactoring~\cite{Mens:2004}, another popular way to improve software quality,
is also inadequate to eliminate this error.  A standard way to fix such
invalid thread access error is to replace direct GUI object accessing operations
with asynchronized message passing, to ensure GUI object are accessed in
the UI thread. For example, fixing the bug in Figure~\ref{fig:androiderror}
needs to wrap line 222 inside a \CodeIn{post} message passing method\footnote{The
\CodeIn{post} method is in class \CodeIn{android.widget.View}, providing
a standard utility to send messages to the UI-thread}, as follows:

@@

One may wonder that a straightforward program transformation to wrap
every possible UI-accessing operation with asynchronized message passing
may solve this problem. However, such an approach is problematic,
because using asynchronized message passing has no timing
guarantee. It is entirely possible that, in some cases, a GUI object has already been
disposed before the message sent to arrives. Thus, programmers can not use
it everywhere; instead, they must choose carefully to put it in appropriate places
with extreme caution.

@@

In this paper, we choose to use static analysis to find such potential errors.
Static analysis has several advantages while comparing to dynamic approaches like testing for this problem.
First, a static analysis can explore every path of the program without executing the code, and
thus there is no need to construct non-trivial tests and harnesses.
Second, a static analysis can soundly find potentially errors: if it
reports no warnings, the code is guaranteed to be bug-free.

@@

\item \textbf{Technique.} We present a general technique to find
such errors. In particular, to deal with the unique features in
multithreaded GUI programs, we propose a combined RTA and k-CFA
call graph construction algorithm

\item \textbf{Implementation.} We instantiate our technique for four
popular GUI platforms: SWT, eclipse plugin, Swing, and Android. Our
tool implementation is publicly available at:
\url{http://guierrordetector.googlecode.com}

### 2011-12-22 GUIErrorDetector/paper/technique.tex
tweak

\subsection{Problem Formulation}

formulate it as a graph reachability

@@

>>>>

We first gives our high-level approach formulation, then present
an error detection algorithm. Finally, we show how to instantiate
the error detection to different GUI frameworks.

\subsection{Approach Formulation}

This section formulates our high-level approach. We define the
notations of \textit{UI thread}, \textit{non-UI threads},
\textit{UI accessing methods},
 \textit{Safe methods}, and \textit{invalid thread access error}, and
two assumptions we make with regard to error detection.

\noindent {\textsc{\textbf{Definition 1 (UI Thread).}}} {UI thread
is a special thread created by the GUI framework during
GUI initialization. After the GUI becomes visible, the UI thread
takes charge of the application, handle events from the GUIs,
or spawn new threads to process lengthy operations in the background. }\vspace{1mm}

\noindent {\textsc{\textbf{Assumption 1.}}} {We assume that each multithreaded
GUI application has a single global UI thread. This is true for applications
built on top of GUI frameworks that adapt the single-thread rule. It is
possible that an application launches another application with its own UI thread
by initializing a different process. In that case, we require the new
launched application to be analyzed separately.}\vspace{1mm}

\noindent {\textsc{\textbf{Definition 2 (non-UI Thread).}}} {Any other
threads except for the UI thread is called non-UI threads.}\vspace{1mm}

\noindent {\textsc{\textbf{Assumption 2.}}} { We assume that each non-UI
thread can only be spawned by the UI thread. Under this assumption, all post-initialization
GUI work naturally occurs in the UI-thread. Once the GUI is visible, the
application is driven by events, which are always handled in the UI thread. We
believe this assumption is reasonable, since if a non-UI
thread is spawned during pre-initialization GUI work and that thread accesses
a GUI object, an exception becomes apparent and the whole application may abort. This
is unlikely for most of the realistic GUI applications.
}\vspace{1mm}

\noindent {\textsc{\textbf{Definition 3 (UI Accessing Methods).}}} { A method
whose execution may read or write a UI object is called a UI Accessing Method.}\vspace{1mm}

\noindent  {\textsc{\textbf{Definition 4 (Safe Methods).}}} {GUI frameworks that
adapt the single-thread rule must provide methods to permit non-UI threads
to run code in the UI thread. We can such methods \textit{safe methods}, since
they can be invoked safely by any threads.}\vspace{1mm}

\noindent  {\textsc{\textbf{Definition 5 (Invalid Thread Access).}}} {An invalid
thread access error occurs if the UI thread spawns a non-UI thread, and there
exists a path from the non-UI thread's \CodeIn{start} method to any UI accessing methods
without going through any safe methods. }\vspace{2mm}

At a high level, to detect an invalid thread access error, an analysis only needs to track all
non-UI threads spawned by the UI thread, and check whether those non-UI threads
may invoke a UI accessing method.

@@

We cast the problem of finding invalid thread

Choose a right abstraction, why call graph is good?

a call graph is a good approximation

Given a sound call graph, our algorithm is sound in that it does not
miss true positives. We note, however, that coomputing a sound
call graph for partial programs such as an eclipse plugin,
or in the presence of reflection, is non-trivial. Existing
call graph construction algorithms either suffer from incompleteness
or imprecision.  XX We describe our combined xxx in Section XXX.

We also note that static analysis may report redundant
